\documentclass[draft=false
              ,paper=a4
              ,twoside=false
              ,fontsize=11pt
              ,headsepline
              ,BCOR10mm
              ,DIV11
              ]{scrbook}
\usepackage[ngerman,english]{babel}
%% see http://www.tex.ac.uk/cgi-bin/texfaq2html?label=uselmfonts
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{libertine}
\usepackage{pifont}
\usepackage{microtype}
\usepackage{textcomp}
\usepackage[german,refpage]{nomencl}
\usepackage{setspace}
\usepackage{makeidx}
\usepackage{listings}
\usepackage{natbib}
\usepackage[ngerman,colorlinks=true]{hyperref}
\usepackage{soul}
\usepackage[printer]{hawstyle}

%% define some colors
\colorlet{BackgroundColor}{gray!20}
\colorlet{KeywordColor}{blue}
\colorlet{CommentColor}{black!60}
%% for tables
\colorlet{HeadColor}{gray!60}
\colorlet{Color1}{blue!10}
\colorlet{Color2}{white}

%% configure colors
\HAWifprinter{
  \colorlet{BackgroundColor}{gray!20}
  \colorlet{KeywordColor}{black}
  \colorlet{CommentColor}{gray}
  % for tables
  \colorlet{HeadColor}{gray!60}
  \colorlet{Color1}{gray!40}
  \colorlet{Color2}{white}
}{}
\lstset{%
  numbers=left,
  numberstyle=\tiny,
  stepnumber=1,
  numbersep=5pt,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{KeywordColor}\bfseries,
  identifierstyle=\color{black},
  commentstyle=\color{CommentColor},
  backgroundcolor=\color{BackgroundColor},
  captionpos=b,
  fontadjust=true
}
\lstset{escapeinside={(*@}{@*)}, % used to enter latex code inside listings
        morekeywords={uint32_t, int32_t}
}
\ifpdfoutput{
  \hypersetup{bookmarksopen=false,bookmarksnumbered,linktocpage}
}{}

%% more fancy C++
\DeclareRobustCommand{\cxx}{C\raisebox{0.25ex}{{\scriptsize +\kern-0.25ex +}}}

\clubpenalty=10000
\widowpenalty=10000
\displaywidowpenalty=10000

% unknown hyphenations
\hyphenation{
}

%% recalculate text area
\typearea[current]{last}

\makeindex
\makenomenclature

\begin{document}
\selectlanguage{english}

%%%%%
%% customize (see readme.pdf for supported values)
\HAWThesisProperties{Author={Jakob Joachim}
                    ,Title={Methodology for Debugging Flink Applications}
                    ,EnglishTitle={Methodology for Debugging Flink Applications}
                    ,ThesisType={Bachelorarbeit}
                    ,ExaminationType={Bachelorprüfung}
                    ,DegreeProgramme={Bachelor of Science Angewandte Informatik}
                    ,ThesisExperts={Prof. Dr. Olaf Zukunft \and Prof. Dr. Klaus-Peter Kossakowski}
                    ,ReleaseDate={2. February 2018}
                    }
%% title
\frontmatter

%% output title page
\maketitle

\onehalfspacing

%% add abstract pages
%% note: this is one command on multiple lines
\HAWAbstractPage
%% German abstract
{Flink, Debugging, Stream Processing, Methodik zum Debuggen, Java Debuggen, Flink Debuggen}%
{Das Debuggen von Programmen ist kompliziert, wird aber in einem verteilten System noch komplizierter. Je größer die täglich erzeugten Datenmengen werden, desto wichtiger wird die Nutzung verteilter Anwendung. Flink ist ein Framework zur verteilten Verarbeitung von Daten aus einem Stream, wobei die Daten zwischen den einzelnen Verarbeitungsschritten ebenfalls durch Streams transportiert werden. Flink erleichtert es dem Entwickler, verteilte Anwendungen, die auf der Datenverarbeitung basieren, zu entwickeln. Diese Thesis stellt eine Methodik zum Debuggen von eben diesen Flink Anwendungen dar.}
%% English abstract
{Flink, Debugging, Stream Processing, Methodology for Debugging, Debugging Java, Debugging Flink}%
{Debugging Applications is hard, debugging distributed applications is harder and the more massive the amount of data that gets generated every day gets, the higher the need for these distributed applications. Flink is a distributed stream processing framework; it processes data while only using streams. It makes it easy for developers to write these distributed applications to handle the ever-growing amounts of data in the world. This thesis explains how to debug these Flink applications by providing a debugging methodology.}

\newpage
\singlespacing

\tableofcontents
\newpage
%% enable if these lists should be shown on their own page
%%\listoftables
%%\listoffigures
\lstlistoflistings
\listoffigures

%% main
\mainmatter
\onehalfspacing
%% write to the log/stdout
\typeout{===== File: chapter 1}
\chapter{Introduction}
  \input{introduction.tex}

\typeout{===== File: chapter 2}
\chapter{The Art of Debugging}
  \input{theArtOfDebugging.tex}

\typeout{===== File: chapter 3}
\chapter{Related Work}
  \input{relatedWork.tex}

\typeout{===== File: chapter 4}
\chapter{Debugging Flink Applications}
  \input{methodForDebuggingFlink.tex}

\typeout{===== File: chapter 5}
\chapter{Flink Backtracker}
\label{flinkBacktracker}
  \input{flinkBacktracker.tex}

\typeout{===== File: chapter 6}
\chapter{Conclusion and Outlook}
  \input{conclusion.tex}

%% appendix if used
%%\appendix
%%\typeout{===== File: appendix}
%%\include{appendix}

% bibliography and other stuff
\backmatter

\typeout{===== Section: literature}
%% read the documentation for customizing the style
\bibliographystyle{dinat}
\bibliography{thesis}

\typeout{===== Section: nomenclature}
%% uncomment if a TOC entry is needed
%%\addcontentsline{toc}{chapter}{Glossar}
\renewcommand{\nomname}{Glossar}
\clearpage
\markboth{\nomname}{\nomname} %% see nomencl doc, page 9, section 4.1
\printnomenclature

%% index
\typeout{===== Section: index}
\printindex

\HAWasurency

\end{document}
